19Q1

### 2019-05-02 19Q1.Rmd
updated Rmd

```{r}
library(circlize)
library(ComplexHeatmap)
library(cowplot)
library(crunch)
library(ggpubr)
library(ggrepel)
library(ggsignif)
library(kableExtra)
library(naniar)
library(openxlsx)
library(reshape2)
library(tidyverse)
library(VennDiagram)

@@

# Import data

@@

This comprehensive cancer cell line information was curated by both Daniel Charytonowicz and myself.
```{r}
ccl_converter <- read.delim("./data_munging/cell_line_database_v1_20180911.tsv", row.names = 1, sep = "\t", header = TRUE)
lineage_converter <- unique(ccl_converter[, c("Broad_ID", "grouped_general_doid", "group_general_lineage_name")])
ccl_histologies <- read.delim("./data_munging/ccl_histologies_metadata.csv", sep = ",", header = TRUE, row.names = 1)
```

@@

DepMap 19Q1 update fixed almost all of the inconsistencies of the Primary.Disease column, though there were come capitalization inconsistencies.
```{r}
ccl_meta <- read.delim("./data_munging/19Q1/DepMap-2019q1-celllines_v2_20190305.csv", sep = ",", header = TRUE, na.strings = c("", NA))
ccl_meta$Primary.Disease <- tolower(ccl_meta$Primary.Disease)
ccl_meta$Primary.Disease <- with(ccl_meta, ifelse(Primary.Disease == "colon/colorectal cancer", "colorectal cancer", Primary.Disease))
ccl_meta <- subset(ccl_meta, Primary.Disease != "unknown" & Primary.Disease != "primary cells" & !grepl("MERGED", CCLE_Name, fixed = TRUE))
ccl_meta$Cell_line <- as.character(lapply(ccl_meta$CCLE_Name, function(x) strsplit(as.character(x), split = "_")[[1]][1]))
ccl_meta$Histology <- tolower(gsub("^[a-z0-9]*_", "", ccl_meta$CCLE_Name, ignore.case = TRUE))
# Convert Primary.Disease entries to all lower case to fix factorization issue
ccl_meta$Primary.Disease <- as.factor(tolower(ccl_meta$Primary.Disease))
# Harmonize ID column name
colnames(ccl_meta)[colnames(ccl_meta) == "DepMap_ID"] <- "Broad_ID"

@@

```{r}
g2p_all <- read.delim("./data_munging/g2p_full_level_A_dataset_NEW_DEC_2_2018_withGroupedLineages.csv", sep = "\t", header = TRUE)
g2p_all[, 1:2] <- NULL
g2p_all$Drug <- tolower(g2p_all$Drug)
g2p_all[g2p_all == -1] <- NA
g2p_all$Ref[g2p_all$Ref == "None"] <- NA
g2p_all$Alt[g2p_all$Alt == "None"] <- NA
g2p_all <- subset(g2p_all, grepl("^CID", g2p_all$DrugID))
g2p <- merge(g2p_all, lineage_converter, by = "grouped_general_doid", all.x = TRUE)

g2p$Drug_Gene <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene), NA, paste0(g2p$Drug, "_", g2p$Gene))
g2p$Drug_Gene_Lin <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene) | is.na(g2p$group_general_lineage_name), NA, paste0(g2p$Drug, "_", g2p$Gene, "_", g2p$group_general_lineage_name))
g2p$position <- paste0("g.chr", g2p$Chromosome, ":", g2p$End, g2p$Ref, ">", g2p$Alt)
g2p$position <- ifelse(grepl("NA", g2p$position), NA, g2p$position)

@@

```{r, eval=FALSE}
unique(g2p$group_general_lineage_name)
levels(g2p$group_general_lineage_name)

g2p_nonCID <- subset(g2p_all, !grepl("^CID", g2p_all$DrugID))
summ <- g2p_nonCID %>% group_by(Drug, DrugID) %>% tally() %>% summarize(DrugID = DrugID, n = n, percent = formatC(n / 402 * 100, digits = 6, format = "f"))
summ
```

@@

Mutation calls (coding region, germline filtered):
```{r, eval=FALSE}
maf_raw <- read.delim("./data_munging/19Q1/depmap_19Q1_mutation_calls_20190305.csv.gz", header = TRUE, sep = ",", row.names = 1)
# Harmonize ID column name
colnames(maf_raw)[colnames(maf_raw) == "DepMap_ID"] <- "Broad_ID"
colnames(maf_raw)[colnames(maf_raw) == "Hugo_Symbol"] <- "Gene"
# Select columns
maf_df <- subset(maf_raw, select = c("Gene", "Broad_ID", "Variant_Classification", "Reference_Allele", "Tumor_Seq_Allele1"))
maf_df$Reference_Allele <- ifelse(maf_df$Reference_Allele == "-", 0, nchar(as.character(maf_df$Reference_Allele)))
maf_df$Tumor_Seq_Allele1 <- nchar(as.character(maf_df$Tumor_Seq_Allele1))
maf_df <- maf_df[maf_df$Variant_Classification != "Silent",]
# Add Mutation_Status column
maf_df$Mutation_Status <- 1
maf_df$PointMutation <- with(maf_df, ifelse(Reference_Allele == 1 | is.na(Reference_Allele), TRUE, ifelse(Reference_Allele == 0 & Tumor_Seq_Allele1 == 1, 1, 0)))

@@

DepMap WES CN Data:
```{r, eval=FALSE}
cn <- read.delim("./data_munging/19Q1/public_19Q1_gene_cn_20190305.csv.gz", sep = ",", check.names = FALSE, header = TRUE)
# Remove Entrez gene IDs from colnames, format is "Gene (Entrez ID)"
colnames(cn) <- gsub(" .*", "", colnames(cn))
colnames(cn)[1] <- "Broad_ID"
# Melt
cn_melt <- melt(data = cn, id.vars = "Broad_ID", measure.vars = colnames(cn[2:ncol(cn)]), variable.name = "Gene", value.name = "CN")

@@

19Q1, CCLE RNAseq gene expression data (log2(TPM+1)):
```{r, eval=FALSE}
ge <- read.delim("./../htscreens_giant_files/CCLE_depMap_19Q1_TPM_20190305.csv.gz", header = TRUE, sep = ",", check.names = FALSE)
# Remove Entrez gene IDs from colnames, format is "Gene (Entrez ID)"
colnames(ge) <- gsub(" .*", "", colnames(ge_19Q1))
colnames(ge)[1] <- "Broad_ID"
# Melt
ge_melt <- melt(ge, id.vars = "Broad_ID", measure.vars = colnames(ge_19Q1)[2:ncol(ge_19Q1)], variable.name = "Gene", value.name = "TPM")

@@

Load Rds:
```{r}
maf_df <- readRDS("./data_munging/19Q1/maf_df_19Q1.rds")
cn_melt <- readRDS("./data_munging/19Q1/cn_melt_19Q1.rds")
ge_melt <- readRDS("./data_munging/19Q1/ge_melt_19Q1.rds")
```

# Merge omics

@@

```{r}
test <- filter(maf_df, Gene %in% g2p_genes) %>% count(Gene, Broad_ID)
test <- test[order(test$n, decreasing = TRUE),]
```

`r nrow(filter(test, n != 1))`/`r nrow(test)` gene-cell line combinations have more than one mutation in the MAF file. This is corrected in the point mutation filtering done below.
```{r, eval=FALSE}
# Load in Pozdeyev et al. (2016) data
ccle_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CCLE_drug_response_metrics.csv.gz", sep = ",")
ctrp_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CTRP_drug_response_metrics.csv.gz", sep = ",")
gdsc_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/GDSC_drug_response_metrics.csv.gz", sep = ",")

@@

```{r, eval=TRUE}
ccle <- readRDS("./data_munging/rds/ccle_data_19Q1_g2pgenes.rds")
ctrp <- readRDS("./../htscreens_giant_files/ctrp_data_19Q1_g2pgenes.rds")
gdsc <- readRDS("./data_munging/rds/gdsc_data_19Q1_g2pgenes.rds")
ccle_ptmuts <- readRDS("./data_munging/rds/ccle_data_19Q1_g2pgenes_ptmuts.rds")
ctrp_ptmuts <- readRDS("./../htscreens_giant_files/ctrp_data_19Q1_g2pgenes_ptmuts.rds")
gdsc_ptmuts <- readRDS("./data_munging/rds/gdsc_data_19Q1_g2pgenes_ptmuts.rds")

@@

How many datasets screen each G2P drug?
```{r, eval=FALSE}
drug_summ <- data.frame(Drug = character(), Dataset_Count = integer(), stringsAsFactors = FALSE)

@@

Merge all CRISPR genes:
```{r, eval=FALSE}
crispr_all <- read.delim("./data_munging/19Q1/gene_effect_corrected_20190305.csv.gz", sep = ",", header  = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(crispr_all) <- gsub(" .*", "", colnames(crispr_all))
colnames(crispr_all)[1] <- "Broad_ID"

@@

```{r, eval=TRUE}
crispr_data <- readRDS("./../htscreens_giant_files/crispr_data_19Q1.rds")

# Filter G2P CRISPR for point mutations
crispr_ptmuts <- filter(crispr_data, PointMutation == 1)
crispr_ptmuts <- subset(crispr_ptmuts, select = -c(PointMutation))
crispr_ptmuts <- filter(crispr_ptmuts, Gene %in% g2p_genes)
```

# G2P summary heatmap

@@

```{r fig.height=6, fig.width=8}
g2p_dg_counts <- g2p_all %>% count(Drug, Gene) %>% filter(!is.na(Drug), !is.na(Gene))
g2p_dg_counts <- dcast(g2p_dg_counts, Drug ~ Gene, value.var = "n", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Drug)
g2p_dg_counts <- g2p_dg_counts[order(rowSums(-g2p_dg_counts)), order(colSums(-g2p_dg_counts))]
g2p_dg_counts[g2p_dg_counts == 0] <- NA

@@

# Wilcoxon tests (mut)

@@

WilcoxonTestDF <- function(dataframe, measure_var, grouping, outfile) {
  require(ggpubr)
  signif <- compare_means(Score ~ Mutation_Status, group.by = grouping, data = dataframe, method = "wilcox.test", p.adjust.method = "BH")
  signif <- adj_signif(signif, alpha = 0.05)
  signif <- signif[order(signif$p),]
  saveRDS(signif, outfile)
  return(signif)
}
```

```{r, eval=FALSE}
crispr_signif_gene <- WilcoxonTestDF(dataframe = crispr_g2p, grouping = "Gene", outfile = "./data_munging/rds/crispr_signif_g2p_gene.rds")
crispr_signif_lineage <- WilcoxonTestDF(dataframe = crispr_g2p, grouping = c("Gene", "Primary.Disease"), outfile = "./data_munging/rds/crispr_signif_g2p_lineage.rds")

ccle_signif_gene <- WilcoxonTestDF(dataframe = ccle_ptmuts, grouping = "Drug_Gene", outfile = "./data_munging/rds/ccle_signif_g2p_gene.rds")
ccle_signif_gene$G2P_Drug_Gene <- ifelse(ccle_signif_gene$Drug_Gene %in% g2p_druggenes, TRUE, FALSE)
ccle_signif_lineage <- WilcoxonTestDF(dataframe = filter(ccle_ptmuts, Drug_Gene_Lin %in% g2p_dgls), grouping = "Drug_Gene_Lin", outfile = "./data_munging/rds/ccle_signif_g2p_lineage.rds")

ctrp_signif_gene <- WilcoxonTestDF(dataframe = ctrp_ptmuts, grouping = "Drug_Gene", outfile = "./data_munging/rds/ctrp_signif_g2p_gene.rds")
ctrp_signif_gene$G2P_Drug_Gene <- ifelse(ctrp_signif_gene$Drug_Gene %in% g2p_druggenes, TRUE, FALSE)
ctrp_signif_lineage <- WilcoxonTestDF(dataframe = filter(ctrp_ptmuts, Drug_Gene_Lin %in% g2p_dgls), grouping = "Drug_Gene_Lin", outfile = "./data_munging/rds/ctrp_signif_g2p_lineage.rds")

gdsc_signif_gene <- WilcoxonTestDF(dataframe = gdsc_ptmuts, grouping = "Drug_Gene", outfile = "./data_munging/rds/gdsc_signif_g2p_gene.rds")
gdsc_signif_gene$G2P_Drug_Gene <- ifelse(gdsc_signif_gene$Drug_Gene %in% g2p_druggenes, TRUE, FALSE)
gdsc_signif_lineage <- WilcoxonTestDF(dataframe = filter(gdsc_ptmuts, Drug_Gene_Lin %in% g2p_dgls), grouping = "Drug_Gene_Lin", outfile = "./data_munging/rds/gdsc_signif_g2p_lineage.rds")
```

```{r}
crispr_signif_gene <- readRDS("./data_munging/rds/crispr_signif_g2p_gene.rds")
crispr_signif_lineage <- readRDS("./data_munging/rds/crispr_signif_g2p_lineage.rds")
ccle_signif_gene <- readRDS("./data_munging/rds/ccle_signif_g2p_gene.rds")
ccle_signif_lineage <- readRDS("./data_munging/rds/ccle_signif_g2p_lineage.rds")
ctrp_signif_gene <- readRDS("./data_munging/rds/ctrp_signif_g2p_gene.rds")
ctrp_signif_lineage <- readRDS("./data_munging/rds/ctrp_signif_g2p_lineage.rds")
gdsc_signif_gene <- readRDS("./data_munging/rds/gdsc_signif_g2p_gene.rds")
gdsc_signif_lineage <- readRDS("./data_munging/rds/gdsc_signif_g2p_lineage.rds")
```

@@

# Spearman (CN, GE)

@@

Spearman correlations for gene expression and copy number:
```{r, warning=FALSE}
crispr_spear_ge <- transform(crispr_ptmuts %>% group_by(Gene) %>% summarize(p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value, p_log10 = -log10(p), corr = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate, Metric = "Gene Expression"), Gene = reorder(Gene, corr))
crispr_spear_cn <- transform(crispr_ptmuts %>% group_by(Gene) %>% summarize(p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value, p_log10 = -log10(p), corr = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate, Metric = "Copy Number"), Gene = reorder(Gene, corr))
```

Create CRISPR two-column, KRAS-inset plot for CN and GE:
```{r fig.height=6, fig.width=12}
spearman_KRAS <- rbind(filter(crispr_ptmuts, Gene == "KRAS") %>% summarize(p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value, corr = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate, Metric = "Gene Expression"), filter(crispr_ptmuts, Gene == "KRAS") %>% summarize(p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value, corr = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate, Metric = "Copy Number"))

>>>>

```{r libraries}
library(circlize)
library(ComplexHeatmap)
library(cowplot)
library(crunch)
library(ggpubr)
library(ggrepel)
library(ggsignif)
library(kableExtra)
library(naniar)
library(openxlsx)
library(reshape2)
library(scales)
library(tidyverse)
library(VennDiagram)

@@

# DATA

@@

This comprehensive cancer cell line information was curated by both Daniel Charytonowicz and myself.
```{r dan_ccl_database}
ccl_converter <- read.delim("./data_munging/cell_line_database_v1_20180911.tsv", row.names = 1, sep = "\t", header = TRUE)
ccl_lineage_converter <- unique(ccl_converter[, c("Broad_ID", "grouped_general_doid", "group_general_lineage_name")])
lineage_converter <- unique(select(lineage_converter, grouped_general_doid, group_general_lineage_name))
ccl_histologies <- read.delim("./data_munging/ccl_histologies_metadata.csv", sep = ",", header = TRUE, row.names = 1)
```

@@

DepMap 19Q1 update fixed almost all of the inconsistencies of the Primary.Disease column, though there were come capitalization inconsistencies.
```{r depmap_metadata}
ccl_meta <- read.delim("./data_munging/19Q1/DepMap-2019q1-celllines_v2_20190305.csv", sep = ",", header = TRUE, na.strings = c("", NA))
ccl_meta$Primary.Disease <- tolower(ccl_meta$Primary.Disease)
ccl_meta$Primary.Disease <- with(ccl_meta, ifelse(Primary.Disease == "colon/colorectal cancer", "colorectal cancer", Primary.Disease))
ccl_meta <- subset(ccl_meta, Primary.Disease != "unknown" & Primary.Disease != "primary cells" & !grepl("MERGED", CCLE_Name, fixed = TRUE))
ccl_meta$Cell_line <- as.character(lapply(ccl_meta$CCLE_Name, function(x) strsplit(as.character(x), split = "_")[[1]][1]))
ccl_meta$Histology <- tolower(gsub("^[a-z0-9]*_", "", ccl_meta$CCLE_Name, ignore.case = TRUE))
# Convert Primary.Disease entries to all lower case to fix factorization issue
ccl_meta$Primary.Disease <- as.factor(tolower(ccl_meta$Primary.Disease))
# Harmonize ID column name
colnames(ccl_meta)[colnames(ccl_meta) == "DepMap_ID"] <- "Broad_ID"

@@

```{r g2p_interpretations}
g2p <- read.delim("./data_munging/g2p_full_level_A_dataset_NEW_DEC_2_2018_withGroupedLineages.csv", sep = "\t", header = TRUE)
g2p[, 1:2] <- NULL
g2p$Drug <- tolower(g2p$Drug)
g2p[g2p == -1] <- NA
g2p$Ref[g2p$Ref == "None"] <- NA
g2p$Alt[g2p$Alt == "None"] <- NA
g2p <- subset(g2p, grepl("^CID", g2p$DrugID))
g2p$Drug_Gene <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene), NA, paste0(g2p$Drug, "_", g2p$Gene))
g2p$position <- paste0("g.chr", g2p$Chromosome, ":", g2p$End, g2p$Ref, ">", g2p$Alt)
g2p$position <- ifelse(grepl("NA", g2p$position), NA, g2p$position)
g2p <- merge(g2p, lineage_converter, by = "grouped_general_doid", all.x = TRUE)
g2p$Drug_Gene_Lin <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene) | is.na(g2p$group_general_lineage_name), NA, paste0(g2p$Drug, "_", g2p$Gene, "_", g2p$group_general_lineage_name))

@@

```{r testing, eval=FALSE}
unique(g2p$group_general_lineage_name)
levels(g2p$group_general_lineage_name)

g2p_nonCID <- subset(g2p, !grepl("^CID", g2p$DrugID))
summ <- g2p_nonCID %>% group_by(Drug, DrugID) %>% tally() %>% summarize(DrugID = DrugID, n = n, percent = formatC(n / 402 * 100, digits = 6, format = "f"))
summ
```

@@

Mutation calls (coding region, germline filtered):
```{r mutation_status, eval=FALSE}
maf_raw <- read.delim("./data_munging/19Q1/depmap_19Q1_mutation_calls_20190305.csv.gz", header = TRUE, sep = ",", row.names = 1)
# Harmonize ID column name
colnames(maf_raw)[colnames(maf_raw) == "DepMap_ID"] <- "Broad_ID"
colnames(maf_raw)[colnames(maf_raw) == "Hugo_Symbol"] <- "Gene"
# Select columns
maf_df <- subset(maf_raw, select = c("Gene", "Broad_ID", "Variant_Classification", "Reference_Allele", "Tumor_Seq_Allele1"))
maf_df$Reference_Allele <- ifelse(maf_df$Reference_Allele == "-", 0, nchar(as.character(maf_df$Reference_Allele)))
maf_df$Tumor_Seq_Allele1 <- nchar(as.character(maf_df$Tumor_Seq_Allele1))
maf_df <- maf_df[maf_df$Variant_Classification != "Silent",]
# Add Mutation_Status column
maf_df$Mutation_Status <- 1
maf_df$PointMutation <- with(maf_df, ifelse(Reference_Allele == 1 | is.na(Reference_Allele), TRUE, ifelse(Reference_Allele == 0 & Tumor_Seq_Allele1 == 1, 1, 0)))

@@

DepMap WES CN Data:
```{r copy_number, eval=FALSE}
cn <- read.delim("./data_munging/19Q1/public_19Q1_gene_cn_20190305.csv.gz", sep = ",", check.names = FALSE, header = TRUE)
# Remove Entrez gene IDs from colnames, format is "Gene (Entrez ID)"
colnames(cn) <- gsub(" .*", "", colnames(cn))
colnames(cn)[1] <- "Broad_ID"
# Melt
cn_melt <- melt(data = cn, id.vars = "Broad_ID", measure.vars = colnames(cn[2:ncol(cn)]), variable.name = "Gene", value.name = "CN")

@@

19Q1, CCLE RNAseq gene expression data (log2(TPM+1)):
```{r gene_expression, eval=FALSE}
ge <- read.delim("./../htscreens_giant_files/CCLE_depMap_19Q1_TPM_20190305.csv.gz", header = TRUE, sep = ",", check.names = FALSE)
# Remove Entrez gene IDs from colnames, format is "Gene (Entrez ID)"
colnames(ge) <- gsub(" .*", "", colnames(ge_19Q1))
colnames(ge)[1] <- "Broad_ID"
# Melt
ge_melt <- melt(ge, id.vars = "Broad_ID", measure.vars = colnames(ge_19Q1)[2:ncol(ge_19Q1)], variable.name = "Gene", value.name = "TPM")

@@

Load Rds:
```{r load_omics_rds}
maf_df <- readRDS("./data_munging/19Q1/maf_df_19Q1.rds")
cn_melt <- readRDS("./data_munging/19Q1/cn_melt_19Q1.rds")
ge_melt <- readRDS("./data_munging/19Q1/ge_melt_19Q1.rds")
```

```{r omics_testing, eval=FALSE}
test <- filter(maf_df, Gene %in% g2p_genes) %>% count(Gene, Broad_ID)
test <- test[order(test$n, decreasing = TRUE),]
```

# MERGE OMICS

@@

`r nrow(filter(test, n != 1))`/`r nrow(test)` gene-cell line combinations have more than one mutation in the MAF file. This is corrected in the point mutation filtering done below.
```{r merge_filter_drug_data, eval=FALSE}
# Load in Pozdeyev et al. (2016) data
ccle_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CCLE_drug_response_metrics.csv.gz", sep = ",")
ctrp_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CTRP_drug_response_metrics.csv.gz", sep = ",")
gdsc_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/GDSC_drug_response_metrics.csv.gz", sep = ",")

@@

```{r load_drug_rds, eval=TRUE}
ccle <- readRDS("./data_munging/rds/ccle_data_19Q1_g2pgenes.rds")
ctrp <- readRDS("./../htscreens_giant_files/ctrp_data_19Q1_g2pgenes.rds")
gdsc <- readRDS("./data_munging/rds/gdsc_data_19Q1_g2pgenes.rds")
ccle_ptmuts <- readRDS("./data_munging/rds/ccle_data_19Q1_g2pgenes_ptmuts.rds")
ctrp_ptmuts <- readRDS("./../htscreens_giant_files/ctrp_data_19Q1_g2pgenes_ptmuts.rds")
gdsc_ptmuts <- readRDS("./data_munging/rds/gdsc_data_19Q1_g2pgenes_ptmuts.rds")

@@

How many datasets screen each G2P drug?
```{r drug_testing, eval=FALSE}
drug_summ <- data.frame(Drug = character(), Dataset_Count = integer(), stringsAsFactors = FALSE)

@@

Merge all CRISPR genes:
```{r merge_crispr_data, eval=FALSE}
crispr_all <- read.delim("./data_munging/19Q1/gene_effect_corrected_20190305.csv.gz", sep = ",", header  = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(crispr_all) <- gsub(" .*", "", colnames(crispr_all))
colnames(crispr_all)[1] <- "Broad_ID"

@@

```{r load_crispr_filter, eval=TRUE}
crispr_data <- readRDS("./../htscreens_giant_files/crispr_data_19Q1.rds")

# Filter G2P CRISPR for point mutations
crispr_ptmuts <- filter(crispr_data, PointMutation == 1)
crispr_ptmuts <- subset(crispr_ptmuts, select = -c(PointMutation))
crispr_ptmuts <- filter(crispr_ptmuts, Gene %in% g2p_genes)
crispr_ptmuts$Gene <- factor(crispr_ptmuts$Gene)
crispr_ptmuts$Primary.Disease <- factor(crispr_ptmuts$Primary.Disease)
```

# DATA SUMMARY FIGURES

@@

## Cell line and drug overlaps

```{r ccl_venn, eval=FALSE}
ccle_ccls <- unique(ccle_ptmuts$Broad_ID)
ctrp_ccls <- unique(ctrp_ptmuts$Broad_ID)
gdsc_ccls <- unique(gdsc_ptmuts$Broad_ID)
crispr_ccls <- unique(crispr_ptmuts$Broad_ID)

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn.diagram(x = list(ccle_ccls, ctrp_ccls, gdsc_ccls, crispr_ccls),
             filename = "./plots/manuscript/venn_ccls.png",
             output = TRUE, imagetype = "png",
             height = 2000, width = 2000, resolution = 500,
             lwd = 1, cex = 1,
             category.names = c("CCLE", "CTRP", "GDSC", "CRISPR"),
             cat.cex = 1.4, cat.fontface = "bold", cat.default.pos = "outer",
             # cat.dist = c(0.06, 0.06, 0.035),
             fill = c("springgreen4", "steelblue4", "turquoise4", "darkorchid"))
```

```{r drug_venn, eval=FALSE}
ccle_drugs <- unique(ccle_ptmuts$Drug)
ctrp_drugs <- unique(ctrp_ptmuts$Drug)
gdsc_drugs <- unique(gdsc_ptmuts$Drug)

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn.diagram(x = list(ccle_drugs, ctrp_drugs, gdsc_drugs),
             filename = "./plots/manuscript/venn_drugs.png",
             output = TRUE, imagetype = "png",
             height = 2000, width = 2000, resolution = 500,
             lwd = 1, cex = 1.4,
             category.names = c("CCLE", "CTRP", "GDSC"),
             cat.cex = 1.4, cat.fontface = "bold", cat.default.pos = "outer",
             cat.dist = c(0.06, 0.06, 0.035),
             fill = c("palegreen3", "slategray3", "paleturquoise3"))
```

**RUN THIS IN TEH EVENING**
```{r, eval=FALSE}
ccle_hists <- unique(select(ccle_ptmuts, Broad_ID, Primary.Disease)) %>% count(Primary.Disease)
ccle_hists$Perc <- ccle_hists$n / sum(ccle_hists$n) * 100
ccle_hists$Dataset <- "CCLE"
ctrp_hists <- unique(select(ctrp_ptmuts, Broad_ID, Primary.Disease)) %>% count(Primary.Disease)
ctrp_hists$Perc <- ctrp_hists$n / sum(ctrp_hists$n) * 100
ctrp_hists$Dataset <- "CTRP"
gdsc_hists <- unique(select(gdsc_ptmuts, Broad_ID, Primary.Disease)) %>% count(Primary.Disease)
gdsc_hists$Perc <- gdsc_hists$n / sum(gdsc_hists$n) * 100
gdsc_hists$Dataset <- "GDSC"

drug_hists <- rbind(ccle_hists, ctrp_hists, gdsc_hists)
drug_hists$Primary.Disease <- as.character(drug_hists$Primary.Disease)

ggplot(drug_hists, aes(x = Dataset, y = n, fill = Primary.Disease)) +
  geom_bar(stat = "identity", color = "black")
```

## G2P summary heatmap

```{r dg_heatmap, fig.height=6, fig.width=8}
g2p_dg_counts <- g2p %>% count(Drug, Gene) %>% filter(!is.na(Drug), !is.na(Gene))
g2p_dg_counts <- dcast(g2p_dg_counts, Drug ~ Gene, value.var = "n", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Drug)
g2p_dg_counts <- g2p_dg_counts[order(rowSums(-g2p_dg_counts)), order(colSums(-g2p_dg_counts))]
g2p_dg_counts[g2p_dg_counts == 0] <- NA

@@

# WILCOXON (MUT)

@@

WilcoxonTestDF <- function(dataframe, grouping, outfile) {
  require(ggpubr)
  if(grouping == "Drug_Gene_Lin") {
    # Filter out DGLs with too few samples in at least one mutation status category
    count <- ctrp_ptmuts %>% count(Drug_Gene_Lin, Mutation_Status)
    count2 <- count %>% count(Drug_Gene_Lin)
    count3 <- merge(count, count2, by = "Drug_Gene_Lin", suffixes = c("_sub3", "_1mut")) %>% filter(n_1mut == 2)
    count3$n_sub3 <- ifelse(count3$n_sub3 >= 3, TRUE, FALSE)
    count4 <- count3 %>% filter(n_sub3) %>% count(Drug_Gene_Lin) %>% filter(n == 2)
    dataframe <- filter(dataframe, Drug_Gene_Lin %in% unique(count4$Drug_Gene_Lin))
    dataframe$Drug_Gene_Lin <- factor(dataframe$Drug_Gene_Lin)
  }
  signif <- compare_means(Score ~ Mutation_Status, group.by = grouping, data = dataframe, paired = FALSE, method = "wilcox.test", p.adjust.method = "BH")
  signif <- adj_signif(signif, alpha = 0.05)
  signif <- signif[order(signif$p),]
  if(grouping == "Drug_Gene") {
    signif$InG2P <- ifelse(signif$Drug_Gene %in% g2p_druggenes, "Yes", "No")
  }
  if(grouping == "Drug_Gene_Lin") {
    signif$InG2P <- ifelse(signif$Drug_Gene_Lin %in% g2p_dgls, "Yes", "No")
  }
  saveRDS(signif, outfile)
  return(signif)
}
```

```{r wilcoxon_tests, eval=FALSE}
crispr_signif_gene <- WilcoxonTestDF(dataframe = crispr_ptmuts, grouping = "Gene", outfile = "./data_munging/rds/crispr_signif_g2p_gene.rds")
crispr_signif_lineage <- WilcoxonTestDF(dataframe = crispr_ptmuts, grouping = c("Gene", "Primary.Disease"), outfile = "./data_munging/rds/crispr_signif_g2p_lineage.rds")

ccle_signif_gene <- WilcoxonTestDF(dataframe = ccle_ptmuts, grouping = "Drug_Gene", outfile = "./data_munging/rds/ccle_signif_g2p_gene.rds")
ccle_signif_lineage <- WilcoxonTestDF(dataframe = ccle_ptmuts, grouping = "Drug_Gene_Lin", outfile = "./data_munging/rds/ccle_signif_g2p_lineage.rds")

ctrp_signif_gene <- WilcoxonTestDF(dataframe = ctrp_ptmuts, grouping = "Drug_Gene", outfile = "./data_munging/rds/ctrp_signif_g2p_gene.rds")
ctrp_signif_lineage <- WilcoxonTestDF(dataframe = ctrp_ptmuts, grouping = "Drug_Gene_Lin", outfile = "./data_munging/rds/ctrp_signif_g2p_lineage.rds")

gdsc_signif_gene <- WilcoxonTestDF(dataframe = gdsc_ptmuts, grouping = "Drug_Gene", outfile = "./data_munging/rds/gdsc_signif_g2p_gene.rds")
gdsc_signif_lineage <- WilcoxonTestDF(dataframe = gdsc_ptmuts, grouping = "Drug_Gene_Lin", outfile = "./data_munging/rds/gdsc_signif_g2p_lineage.rds")
```

```{r load_wilcoxon_rds}
crispr_signif_gene <- readRDS("./data_munging/rds/crispr_signif_g2p_gene.rds")
crispr_signif_lineage <- readRDS("./data_munging/rds/crispr_signif_g2p_lineage.rds")
ccle_signif_gene <- readRDS("./data_munging/rds/ccle_signif_g2p_gene.rds")
ccle_signif_lineage <- readRDS("./data_munging/rds/ccle_signif_g2p_lineage.rds")
ctrp_signif_gene <- readRDS("./data_munging/rds/ctrp_signif_g2p_gene.rds")
ctrp_signif_lineage <- readRDS("./data_munging/rds/ctrp_signif_g2p_lineage.rds")
gdsc_signif_gene <- readRDS("./data_munging/rds/gdsc_signif_g2p_gene.rds")
gdsc_signif_lineage <- readRDS("./data_munging/rds/gdsc_signif_g2p_lineage.rds")
```

@@

# SPEARMAN (CN, GE)

@@

## Drugscreens

```{r drug_spear_calc, warning=FALSE}
g2p_druggene_tally <- g2p %>% group_by(Drug_Gene) %>% tally() %>% drop_na()

# CCLE
ccle_cor_res <- ccle_ptmuts %>% group_by(Drug_Gene) %>%summarize(
                GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate,
                GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value,
                CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate,
                CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value)
ccle_cor_res <- merge(g2p_druggene_tally, ccle_cor_res, by = "Drug_Gene", all.y = TRUE)
ccle_cor_res$n <- ifelse(is.na(ccle_cor_res$n), 0, ccle_cor_res$n)
ccle_cor_res$InG2P <- ifelse(ccle_cor_res$n != 0, "Yes", "No")
ccle_cor_res$Dataset <- "CCLE"

# CTRP
ctrp_cor_res <- ctrp_ptmuts %>% group_by(Drug_Gene) %>%summarize(
                GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate,
                GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value,
                CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate,
                CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value)
ctrp_cor_res <- merge(g2p_druggene_tally, ctrp_cor_res, by = "Drug_Gene", all.y = TRUE)
ctrp_cor_res$n <- ifelse(is.na(ctrp_cor_res$n), 0, ctrp_cor_res$n)
ctrp_cor_res$InG2P <- ifelse(ctrp_cor_res$n != 0, "Yes", "No")
ctrp_cor_res$Dataset <- "CTRP"
